Smart Representations: {R}ationality and Evolution in a Richer Environment

### 2016-06-24 paper/rev1_paperBlinded.tex
paper revision first half

@@

\begin{abstract}
\normalsize
  Standard applications of evolutionary game theory look at a single, fixed game and focus on
  the evolution of behavior for that game alone. Instead, this paper uses tools from
  evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of representing
  a decision situation, paired with a method for choosing an act based on this 
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that compares a number of classic proposals for decision making. The
  main result is that regret-based preference representations can be evolutionarily selected
  for, despite being non-veridical.
\end{abstract}

@@

If agents deal with a rich and variable environment, they have to face many different choice
situations.
Standard evolutionary game models frequently
simplify reality in two ways that are relevant for present purposes. Firstly, the environment
is represented as a fixed \emph{stage game}; secondly, the focus of
evolutionary selection is behavior for that stage game alone. In contrast, some argue for
studying the evolutionary competition of general \emph{choice mechanisms} in a rich and
variable environment
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche,HammStev12}. To show that
there is no tension here, this paper introduces a method for reconciling established notions of
evolutionary game theory with evolutionary selection of choice mechanisms in variable
environments \citep[see, inter alia, related work
by][]{Harley1981:Learning-the-Ev,ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o,OConnor2016:Evolving-to-Gen}.

In full generality, a choice mechanism associates decision situations with action choices. A crucial
part of a choice mechanism is the \emph{subjective representation} of the decision situation,
in particular the manner of forming preferences and beliefs about a possibly uncertain world (see
figure~\ref{fig:ChoiceMechanism} below). This paper asks: which preference and belief
representations are ecologically valuable and lead to high fitness?
The evolution of
preferences has been subject of recent interest in theoretical economics
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. Here, we argue that questions of preference evolution should take
variability in uncertainty representation into account as well. In relation to this, 
the main contributions of this work are essentially three. First, we show that the way agents form beliefs, and the type of uncertainty that they entertain in particular, is a key aspect for the evolutionary success. 
Second, we demonstrate
that if agents have imprecise probabilistic beliefs \citep[e.g.][]{gardsah82,levi74,walley96}, faithful and objective representation in terms of true, real evolutionary fitness can be outperformed by
subjective (e.g., regret-based) preference representations that deviate
from the true fitness that natural selection operates on.
Third, although a few examples along similar lines have recently been introduced (REFERENCES), our approach based on a meta-game model is very uncustomary in the literature, and contributes to widen the methods of evolutionary game theory beyond the traditional single-game framework.

@@

The paper is organized as follows. Section \ref{sec:rati--subj} reviews different views of rational choice and relates the
approach we take here to the existing literature. Sections \ref{sec:subj-ut,bel,dec-rules} and \ref{sec:subj-rep,ev-fit} introduce the
relevant decision-theoretic and game-theoretic notions, while Sections \ref{sec:metagame} and \ref{sec:model} present the model.
%spells out four choice mechanisms on whose
%evolutionary competition we zoom in 
The main results of the paper are in Section~\ref{sec:results}, and Section \ref{sec:extensions} extends the evolutionary analysis by
introducing more choice mechanisms and alternative models of the environment. Finally, Section \ref{sec:conclusion} concludes.

\section{Rationality and Representations}
\label{sec:rati--subj}

@@

If beliefs and preferences are subjective, there is room for \emph{rationalization} or
\emph{redescriptionism} of observable behavior. For example,
\citet{KahnemannTversky1979:Prospect-Theory} famously demonstrated that what appear to be
violations of rationality norms in human decision making can be explained on the assumption
that subjects' beliefs and preferences deviate systematically from the objectively given
parameters in the presented choice task. Similarly for social decision making, including social
preferences such as fairness, altruism or similar, allows us to describe as rational
empirically observed behavior, such as in experimental Prisoner's Dilemmas or public goods
games, that might otherwise appear irrational \citep[e.g.,][]{fehrschmidt99,charrab02}.

The main objection to redescriptionism is that, without additional constraints, the notion of
rationality is likely to collapse, as it seems possible to deem rational almost everything that
is observed, given the freedom to adjust beliefs and preferences at will. \emph{Normativism}
therefore emphasizes that there are many ways in which ascriptions of beliefs and preferences
should be constrained by normative considerations of rationality as well. As for beliefs, where
possible, subjective beliefs should reflect objective chance. 
%Where this is impossible,
%temporal consistency of choices constrains rational subjective beliefs, else exploitation by
%Dutch books is possible. 
As for preferences, in certain situations it may make sense to
postulate on normative grounds that subjective preferences should be oriented towards tracking
objective fitness. For instance, profit maximization seems a necessary requirement for
evolution in a competitive market because only firms behaving according to profit maximization will survive in the long run \citep[e.g.,][]{alch50,Fried53}.

@@

\iffalse

\begin{figure}
  \centering
  \begin{tikzpicture}[node distance = 2.25cm]

  \begin{scope}

    \node[] (environment) {};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = environment, node distance = 2cm] (situation)
    {environment produces random \\ \textbf{decision situation} $D$};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = situation] (representation) 
    {\textbf{subjective representation} of $D$ \\ \textcolor{gray}{[preference, beliefs, \dots]}};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = representation] (action)
    {\textbf{action choice} $a$};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = action,node distance = 1.5cm] (payoff) {\textbf{payoff} for $a$ in $D$};

    \draw[->, thick]
    ([xshift=-1.75cm,yshift=-0.5cm]situation.center)--([xshift=-1.75cm,yshift=0.55cm]representation.center)
    node[midway, right] {subj.~representation scheme $\gamma$};

    \draw[->, thick]
    ([xshift=-1.75cm,yshift=-0.55cm]representation.center)--([xshift=-1.75cm,yshift=0.5cm]action.center)
    node[midway, right] {action selection function $\kappa$};

    \draw[->, thick]
    ([xshift=-1.75cm,yshift=-0.55cm]action.center)--([xshift=-1.75cm,yshift=0.5cm]payoff.center)
    node[midway, right] {};

    \node[minimum height = 0.75cm, text width=6cm, text
    centered, right of = representation, xshift = 2cm, rotate = 90] (mechanism) {choice mechanism};

    \begin{pgfonlayer}{background}
             \node [rounded corners, dashed,
       very thick,draw=black!40,fit={($(situation.south)+(0,-10pt)$) 
                                     ($(mechanism.west)+(+5pt,+50pt)$) 
                                     % ($(action.north)+(0,+10pt)$) 
                                     ($(representation.west)+(-7pt,0)$)
                                   }] (box) {};
    \end{pgfonlayer}

  \end{scope}

\end{tikzpicture}
  \caption{Schematic representation of action choice. A choice mechanism maps decision
    situation $D$ to action choice $a$ by means of a representation of $D$ in terms of a subjective
    preference and a subjective belief.}
  \label{fig:ChoiceMechanism}
\end{figure}

We submit that meta-games are a useful tool for structuring and sharpening conceptual inquiry
in a domain that is notorious for its fuzziness and elusiveness. To show how this approach can
be applied to the benefit of conceptual debate, we consider the evolution of preferences
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. 
%Concretely, we here pit two different ways of
%representing preferences against each other: (i) veridical representations that correctly
%reflect objective fitness, and (ii) non-veridical subjective preferences in terms of regret
%\citep[e.g.,][]{Savage1951:The-theory-of-s,}. 
We show that subjective, non-veridical preference
representations based on a notion of regret can be adaptive and even outperform objective representations of the true evolutionary fitness, when agents, at least on occasion, hold
\emph{imprecise probabilistic beliefs} \citep[e.g.,][]{gilsch89,levi74,gardsah82} and make
decisions based on a security choice rule (here: the maxmin rule).\footnote{Decision-theoretic background and justifications for the use of maxmin and imprecise probabilities is given in Appendix~\ref{sec:impr-prob-beli}.}

The case study under consideration strips the generality of a meta-game approach down to a perspicuous
comparison between four choice mechanisms, summarized in table~\ref{tab:CMs}.
% the cross-product of two different preference representations (veridical vs.~regret) and two
%different belief representations (precise vs.~imprecise). 
\begin{table}[t]
  \centering
  \begin{tabular}{cccc}
  \multicolumn{3}{c}{choice mechanism} \\  \cmidrule(r){1-3}
    \multicolumn{2}{c}{subjective representation} & action selection & corresponding classic rule \\  \cmidrule(r){1-2}
    preference type & belief type  \\ \midrule
    objective fitness & precise & maxmin EU &  Laplace rule \\
    objective fitness & imprecise & maxmin EU & maxmin \\ 
    regret & precise & maxmin EU & Laplace rule \\
    regret & imprecise & maxmin EU & regret minimization \\ 
  \end{tabular}
  \caption{Overview relevant choice mechanisms and the corresponding (behaviorally equivalent)
    classic choice rules.}
  \label{tab:CMs}
\end{table}
\iffalse
This exclusive selection of choice mechanisms is motivated by a number of
considerations. Firstly, the set of conceivable choice mechanisms is vast. Most of that
conceivable vastness is utter nonsense, and some selection of plausible candidates is required in
any case. Secondly, a restriction to only a small number of suitable candidates will
help demonstrate more distinctly how a meta game approach may help structure philosophical
inquiry, which is our main goal. Thirdly, our case study adds to the recent literature on the
evolution of preferences by making evident the possibly non-trivial interaction between
preference and belief representations, all else equal. This is our secondary goal. 
Finally,
\fi 
The
four choice mechanisms,  
%that we look at here
although phrased in terms of preference and belief representations, map onto well-known classic
proposals for rational decision making (see table~\ref{tab:CMs}). In other words, there is also
motivation from historical relevance for our selection. To see this, we first recapitulate the
relevant classics ---to have a handy label, we call them \emph{classic rules}--- and then show
how they relate to our notion of choice mechanism.

The classic literature on rational choice contains a number of proposals for which acts a
player should choose if she is rational. Maximization of expected utility, as
given in Definition~\ref{def:rationality}, is just one of them, possible only if DM holds a precise probabilistic belief.
Here, we consider three others:
(i) maxmin, (ii) Laplace rule, and (iii) regret minimization. 
%All of these apply to games and
%solitary decision situations.
We focus on games first, but definitions carry over straightforwardly to solitary decisions.

A \emph{choice mechanism} is a map from choice situations to action choices that contains an explicit level, in terms of preferences and beliefs, of subjective representation of
the choice situation. Concretely, let us think of a choice mechanism as a pair of functions
($\gamma,\kappa$), where $\gamma$ is a \emph{subjective representation scheme} and $\kappa$ is
an \emph{action selection function} (see figure~\ref{fig:ChoiceMechanism}).

\noindent This means that objective and regret types are behaviorally indistinguishable, as long as both hold precise beliefs. 
%This will be central in understanding later results. 
On the other hand, if paired with imprecise beliefs, different behavioral predictions
arise.

In the following, we assume that agents are maximally uncertain, so that precise types have a
(singleton set containing a) uniform probability measure
$\overline{\mu} \in \Delta(\Acts_{-i})$ and imprecise types entertain all probabilistic beliefs
about the co-players' behavior $\overline{\Gamma} = \Delta(\Acts_{-i})$. In that case, the
considered choice mechanisms are behaviorally equivalent to classic choice rules, as summarized
in table~\ref{tab:CMs}. This means that most of the following results which we interpret as results about the evolutionary fitness of different subjective representations are also relevant for a comparison of the ecological value of classic choice
rules.

\section{Subjective Utilities, Beliefs and Decision Rules} \label{sec:subj-ut,bel,dec-rules}

We view a choice mechanism as the combination of three different things: a \textit{subjective utility} (or preference), a \textit{subjective belief} and a \textit{decision rule}. In general, the agent's action choice will depend both on the agent's utility at different possible outcomes of the choice situation and on the agent's beliefs about the realization of these outcomes. The decision rule then combines the agent's subjective utility and belief, and dictates how the agent should act in case of a choice situation whose consequences are uncertain. Therefore, a decision rule is a function that associates an action choice with the agent's utility and beliefs:
$$ \text{Decision Rule: Utility }\times\text{ Beliefs } \rightarrow \text{ Actions.}  $$
The utility of an agent can be formally expressed by a function $u:W \times A \rightarrow \mathbb{R} $, where $A$ stands for a (finite) set of actions available to the agent and $W$ is a (finite) set of possible states of the world. There are many different ways to describe beliefs, but here we assume that the agent's beliefs are represented in terms of a (possibly singleton) set of probability functions $\Gamma \subseteq \Delta(W) $ over the possible states of the world. In line with this definition, we can consider different decision rules, such as:
\begin{enumerate}

\item \textbf{Maxmin}: $$ a^*(u,\Gamma)= \argmax_{a \in A} \min_{\mu \in \Gamma} \sum_{w\in W} u(w,a)  \mu(w).$$

\item \textbf{Maximax}: $$ a^*(u,\Gamma)= \argmax_{a \in A} \max_{\mu \in \Gamma} \sum_{w\in W} u(w,a)  \mu(w).$$

\item \textbf{Laplace rule}: $$ a^*(u,\Gamma)= \argmax_{a \in A} \sum_{w\in W} \frac{1}{|W|}  u(w,a) .$$

\item \textbf{Expected utility maximization} (for $\Gamma$ singleton): $$ a^*(u,\Gamma)= \argmax_{a \in A} \sum_{w\in W} u(w,a)  \mu(w).$$

@@

It is worth noticing that both maxmin and maximax boil down to expected utility maximization when the set  $\Gamma$ is a singleton, and in turn expected utility maximization reduces to Laplace rule when the belief $\mu$ is a uniform probability over the states.

\section{Subjective Representations and Evolutionary Fitness} \label{sec:subj-rep,ev-fit}

For a choice mechanism to prescribe an action, the decision rule  needs to be given a specific subjective utility and belief as input, as described in the previous section. We then call \textit{subjective representation} of the decision problem the pair $(u,\Gamma)$ of a subjective utility and a subjective belief.
In other words, a subjective representation is a general way of forming preferences and beliefs about the choice situation.

The literature on evolution of preferences (REFERENCES) mainly focuses on the evolutionary selection of subjective utilities uniquely. Instead, it will be clear from our model that different subjective beliefs may have a fundamental role in determining the evolutionary success of subjective utilities, and therefore the interplay between utilities and beliefs needs to be taken into account explicitly.
So, here we
are most interested in the question which subjective representations are better than others
from an evolutionary point of view. In order to investigate this issue we are going to use the machinery of evolutionary game theory.

A game is an interactive decision situation. 
For a given game $G=\langle N, (A_i,\pi^G_i)_{i \in N} \rangle$, let us denote the evolutionary payoff, or \textit{fitness}, of player $i$ by the function $\pi^G_i: \Pi_{i \in N} A_i \rightarrow \mathbb{R}$, where $A_i$ is player $i$'s (finite) set of actions. For simplicity of exposition we assume that
  all games that are played are symmetric two-player games, that is, games such that: $N:=\{1,2\}$, $A_1 = A_2 $, and $\pi^G_1(a_1, a'_2)= \pi^G_2(a'_1, a_2)=:\pi^G(a,a')$.\footnote{Since payoff functions are symmetric, we will simply write $\pi^G(a, a')$ for $\pi^G_1(a_1, a'_2)$ and $A:=A_1=A_2$, as usual. However, notice that all the definitions and results can be extended to more general cases.} Then, the evolutionary fitness of a subjective representation $(u,\Gamma)$ is measured in terms of the fitness of a
choice mechanism that contains it, all else equal. 
Formally, the fitness of a choice mechanism $c$ against choice mechanism $c'$ in a symmetric two-player game $G=\langle \{1,2\}, A, \pi^G \rangle$ is the
expected payoff of $c$ against $c'$ in $G$:

$$
 F_G(c,c')= \pi^G(a_c^*(u^G,\Gamma),a_{c'}^*(u^G,\Gamma)), \footnote{Whenever a choice mechanism would not select a unique
  action, we assume that the player chooses one of the equally optimal actions at random. I.e., $F_G(c,c')= \sum_{a \in a_c^*(u^G,\Gamma)} \sum_{a' \in a_{c'}^*(u^G,\Gamma)} \frac{1}{|a_c^*(u^G,\Gamma)|} \frac{1}{|a_{c'}^*(u^G,\Gamma)|} \pi^G(a,a')$.}
$$
where $a^*_c$ is the decision rule of choice mechanism $c$. Given the game theoretic setting, the subjective utility $u$ is now a function $u^G:A \times A \rightarrow \mathbb{R}$, and the subjective belief $\Gamma$ is a set of probability functions over the co-player's actions, $ \Gamma \subseteq \Delta(A) $.

However, studying the fitness and the evolutionary selection of different choice mechanisms by the use of a fixed, single game may look like a rather restrictive perspective and possibly biased by the choice of the particular game. To avoid these limitations, we extend the standard single-game framework of evolutionary game theory by considering a \textit{class} of possible games that can be played within a population. Although not completely novel (seeREFERENCES), this approach has started being investigated only very recently, and we maintain that it can shed new light on the evolution of behavior, and it can open to much broader perspectives on evolutionary game theory in general.

\section{The Meta-Game} \label{sec:metagame}

The ground for the evolutionary competition of choice mechanisms will consist of a class $\mathcal{G}$ of possible two-player symmetric games, together with a probability measure $P_{\mathcal{G}}(G)$ for the occurrence probability of game $G \in \mathcal{G}$. In more biological terms, the probability measure $P_{\mathcal{G}}$ encodes the statistical
properties of the environment. We then call \textit{meta-game} the tuple $ MG=\langle CM, \mathcal{G}, P_{\mathcal{G}},F \rangle$, where $CM$ is a set of choice mechanisms, $\mathcal{G}$ is a class of possible games, $P_{\mathcal{G}}(G)$ is the probability of game $G$ to occur, and $F:CM \times CM \rightarrow \mathbb{R}$ is the fitness function, defined as:
\begin{align}
  \label{eq:FittnessChoiceMechGamePairwise}
  F(c, c') = \int P_{\mathcal{G}}(G) \  F_G(c,c') \text{ d} G \,.
\end{align}
Hence, $F(c,c')$ determines the evolutionary payoff of choice mechanism $c$ against $c'$ in the meta-game. The set $CM$ is though of as the set of choice mechanisms present within a given population that is playing the games from the class $\mathcal{G}$. Consequently, it is possible to compute the average fitness of $c$ against the population, that is given by:
\begin{align}
  \label{eq:FittnessChoiceMechGame}
  F(c) = \int P_{M}(c') \ F(c,c') \text{ d} c' = \int \int P_{\mathcal{G}}(G) \  P_{M}(c') \  F_G(c,c') \text{ d} c' \text{ d} G  \,,
\end{align}
where $P_{M}(c')$ is the probability of
encountering a co-player with choice mechanism $c'$.

Meta-games are then abstract models for the evolutionary competition between choice
mechanisms in interactive decision making contexts.
Standard notions of evolutionary game theory apply to meta-games as well. A choice
mechanism $c$ is a strict Nash equilibrium if $F(c,c) > F(c',c)$ for all $c'$; it is
evolutionarily stable if for all $c'$ either (i) $F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$
and $F(c,c') > F(c',c')$; it is neutrally stable if for all $c'$ either (i) $F(c,c) > F(c',c)$
or (ii) $F(c,c) = F(c',c)$ and $F(c,c') \ge F(c',c')$
\citep{Maynard-Smith1982:Evolution-and-t}. Similarly, evolutionary dynamics can be applied to
meta-games. We will later turn towards replicator dynamics
\citep{TaylorJonker1978:Evolutionary-St} and replicator mutator dynamics
\citep[e.g.][]{Nowak2006:Evolutionary-Dy}.
%\todo{footnote with caveat about agent-level processes}

\section{The Model} \label{sec:model}

At first, the main subjective utilities that are going to compete against each other for the survival of the fittest are essentially two:
\begin{enumerate}
\item the \textit{objective} utility, defined by: for all $G \in \mathcal{G}$,
$$\text{obj}^G(a,a')=\pi^G(a,a');$$
\item the \textit{regret}, defined by: for all $G \in \mathcal{G}$,
$$\text{reg}^G(a,a') =\pi^G(a,a')- \max_{a''\in A} \pi^G(a'',a').$$
\end{enumerate}

@@

Although a thorough discussion of this issue goes beyond the scope of this work, let us say that these two kinds of belief underlie two different and alternative views on uncertainty. In front of uncertain events, a strict Bayesian will always form a precise belief, specified by a \textit{single} probability $\mu$. In the absence of any information about future uncertain events, the Bayesian would then invoke the \textit{principle of insufficient reason}, and accordingly choose a uniform probability over the possible outcomes. On the contrary, more recently many authors (REFERENCES) argued against the obligation of representing a belief by means of a single probability measure, opposite to the Bayesian prescriptions. They argue instead in favor of a more general account, according to which uncertainty can be unmeasurable, and represented by a (closed and convex) set of probabilities. This line of thought has been originated in decision theory by the famous Ellsberg's paradoxes (REF), and appears extremely relevant
in game-theoretical contexts too. Indeed, in a recent paper \citet{BattCerrMM15} write:

>>>>

\definecolor{Red}{RGB}{178,34,34}
\newcommand{\mf}[1]{\textcolor{Red}{[mf: #1]}}

@@

\begin{abstract}
  \normalsize Standard applications of evolutionary game theory look at a single, fixed game
  and focus on the evolution of behavior for that game alone. Instead, this paper uses tools
  from evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of subjectively
  representing a decision situation, paired with a method for choosing an act based on this
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that shows how subjective preference representations in terms of regret that differ from the
  actual fitness can be evolutionarily advantageous.
\end{abstract}

@@

If agents deal with a rich and variable environment, they have to face many different choice
situations.  Standard evolutionary game models frequently simplify reality in at least two
ways. Firstly, the environment is represented as a fixed \emph{stage game}; secondly, the focus
of evolutionary selection is behavior for that stage game alone. In contrast, some argue for
studying the evolutionary competition of general \emph{choice mechanisms} in a rich and
variable environment
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche,HammStev12}. In
response to this and adding to recent like-minded approaches, this paper introduces a general
meta-game model that conservatively extends the scope of evolutionary game theory to deal with
evolutionary selection of choice mechanisms in variable environments \citep[see, inter alia,
related work
by][]{Harley1981:Learning-the-Ev,ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o,OConnor2016:Evolving-to-Gen}.\mf{please
insert more references here; maybe use a footnote right here to pick out some of the most
important related pieces and say how our contribution is novel and shiny}

A choice mechanism associates decision situations with action choices. A crucial part of a
choice mechanism is the \emph{subjective representation} of the decision situation, in
particular the manner of forming preferences and beliefs about a possibly uncertain world. To
show the usefulness of the meta-game approach, this paper asks: which preference and belief
representations are ecologically valuable and lead to high fitness? The evolution of
preferences has been subject of recent interest in theoretical economics
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. Here, we argue that questions of preference
evolution should take variability in uncertainty representation into account as well. We
demonstrate that if agents have imprecise probabilistic beliefs
\citep[e.g.][]{gardsah82,levi74,walley96}, faithful and objective representation in terms of
true, real evolutionary fitness can be outperformed by subjective (e.g., regret-based)
preference representations that deviate from the true fitness that natural selection operates
on.

@@

The paper is organized as follows. Section \ref{sec:rati--subj} sets the scene by reviewing
different perspectives no rational choice. Section~\ref{sec:subj-ut,bel,dec-rules} then
introduces the meta-game approach. In doing so it covers key notions such as choice mechanism,
decision rules and subjective representations, all with an eye towards the specific application
to the evolution of preferences in Section~\ref{sec:model}. Section~\ref{sec:results} contains
our main results for this application and Section \ref{sec:extensions} discusses some
interesting extensions.

\section{Rationality and Subjective Representations}
\label{sec:rati--subj}

@@

If beliefs and preferences are subjective, there is room for \emph{rationalization} or
\emph{redescriptionism} of observable behavior. For example,
% \citet{KahnemannTversky1979:Prospect-Theory} famously demonstrated that what appear to be
% violations of rationality norms in human decision making can be explained on the assumption
% that subjects' beliefs and preferences deviate systematically from the objectively given
% parameters in the presented choice task. Similarly
in the case of social decision making, including considerations of fairness allows us to
describe as rational empirically observed behavior, such as in experimental Prisoner's Dilemmas
or public goods games, that might otherwise appear irrational
\citep[e.g.,][]{fehrschmidt99,charrab02}.

The main objection to redescriptionism is that, without additional constraints, the notion of
rationality is likely to collapse, as it seems possible to deem rational almost everything that
is observed, given the freedom to adjust beliefs and preferences at will. \emph{Normativism}
therefore emphasizes that there are many ways in which ascriptions of beliefs and preferences
should be constrained by normative considerations of rationality as well: e.g., subjective
beliefs should reflect objective chance where possible; subjective preferences should be
oriented towards tracking objective fitness. For instance, profit maximization seems a
necessary requirement for evolution in a competitive market because only firms behaving
according to profit maximization will survive in the long run \citep[e.g.,][]{alch50,Fried53}.

@@

\section{Choice Mechanisms  \& Meta-Games} 
\label{sec:subj-ut,bel,dec-rules}

We view a choice mechanism as the combination of three different things: a \textit{subjective utility} (or preference), a \textit{subjective belief} and a \textit{decision rule}. In general, the agent's action choice will depend both on the agent's utility at different possible outcomes of the choice situation and on the agent's beliefs about the realization of these outcomes. The decision rule then combines the agent's subjective utility and belief, and dictates how the agent should act in case of a choice situation whose consequences are uncertain. Therefore, a decision rule is a function that associates an action choice with the agent's utility and beliefs:
$$ \text{Decision Rule: Utility }\times\text{ Beliefs } \rightarrow \text{ Actions.}  $$
The utility of an agent can be formally expressed by a function
$u:W \times A \rightarrow \mathbb{R} $, where $A$ stands for a (finite) set of actions
available to the agent and $W$ is a (finite) set of possible states of the world. There are
many different ways to describe beliefs, but for concreteness of later applications we here
assume that the agent's beliefs are represented in terms of a (possibly singleton) set of
probability functions $\Gamma \subseteq \Delta(W) $ over the possible states of the
world. Examples of well-known decision rules from the literature that will be useful later are:
\begin{enumerate}

\item \textbf{Maxmin}: $$ a^*(u,\Gamma)= \argmax_{a \in A} \min_{\mu \in \Gamma} \sum_{w\in W} u(w,a)  \mu(w)$$

\item \textbf{Maximax}: $$ a^*(u,\Gamma)= \argmax_{a \in A} \max_{\mu \in \Gamma} \sum_{w\in W} u(w,a)  \mu(w)$$

\item \textbf{Laplace rule}: $$ a^*(u,\Gamma)= \argmax_{a \in A} \sum_{w\in W} \frac{1}{|W|}  u(w,a) $$

\item \textbf{Expected utility maximization} (for $\Gamma$ singleton): $$ a^*(u,\Gamma)= \argmax_{a \in A} \sum_{w\in W} u(w,a)  \mu(w)$$

@@

It is worth noticing that both maxmin and maximax boil down to expected utility maximization
when the set $\Gamma$ is a singleton, and in turn expected utility maximization reduces to
Laplace rule when the belief $\mu$ is a uniform probability over the states.  \mf{do we need
  ``maximax''?}

As mentioned previously, for a choice mechanism to prescribe an action, the decision rule needs
to be given a specific subjective utility and belief as input. A \textit{subjective
  representation} of a decision problem is the pair $(u,\Gamma)$, consisting of a subjective
utility and a subjective belief. We are specifically interest in general and systematic ways of
forming such subjective representations across many different decision situations.

% The literature on evolution of preferences (REFERENCES) mainly focuses on the evolutionary
% selection of subjective utilities uniquely. Instead, it will be clear from our model that
% different subjective beliefs may have a fundamental role in determining the evolutionary
% success of subjective utilities, and therefore the interplay between utilities and beliefs
% needs to be taken into account explicitly.  So, here we are most interested in the question
% which subjective representations are better than others from an evolutionary point of
% view. In order to investigate this issue we are going to use the machinery of evolutionary
% game theory.

A game is an interactive decision situation.  For a given game
$G=\langle N, (A_i,\pi^G_i)_{i \in N} \rangle$, let us denote the evolutionary payoff, or
\textit{fitness}, of player $i$ by the function
$\pi^G_i: \Pi_{i \in N} A_i \rightarrow \mathbb{R}$, where $A_i$ is player $i$'s (finite) set
of actions. For simplicity of exposition we assume that all games that are played are symmetric
two-player games where $N:=\{1,2\}$, $A_1 = A_2 $ and
$\pi^G_1(a_1, a'_2)= \pi^G_2(a'_1, a_2)=:\pi^G(a,a')$.\footnote{Since payoff functions are
  symmetric, we will simply write $\pi^G(a, a')$ for $\pi^G_1(a_1, a'_2)$ and $A:=A_1=A_2$, as
  usual. However, notice that all the definitions and results can be extended to more general
  cases.} Then, the evolutionary fitness of a subjective representation $(u_c,\Gamma_c)$ is
measured in terms of the fitness of a choice mechanism that contains it, all else equal.
Formally, the fitness payoff of a choice mechanism $c$ against choice mechanism $c'$ in a
symmetric two-player game $G=\langle \{1,2\}, A, \pi^G \rangle$ is the expected payoff of $c$
against $c'$ in $G$:

$$
F_G(c,c')= \pi^G(a_c^*(u_c^G,\Gamma_c),a_{c'}^*(u_{c'}^G,\Gamma_{c'})), \footnote{Whenever a choice mechanism
  would not select a unique action, we assume that the player chooses one of the equally
  optimal actions at random. I.e., $F_G(c,c')=
  \sum_{a \in a_c^*(u_c^G,\Gamma_c)} \sum_{a' \in a_{c'}^*(u_{c'}^G,\Gamma_{c'})}
  \frac{1}{|a_c^*(u_c^G,\Gamma_c)|} \frac{1}{|a_{c'}^*(u_{c'}^G,\Gamma_{c'})|} \pi^G(a,a')$.}
$$
where $a^*_c$ is the decision rule of choice mechanism $c$. Given the game theoretic setting,
the subjective utility $u_c$ is now a function $u_c^G:A \times A \rightarrow \mathbb{R}$, and
the subjective belief $\Gamma_c$ is a set of probability functions over the co-player's
actions, $ \Gamma \subseteq \Delta(A)$.

However, studying the fitness and the evolutionary selection of different choice mechanisms by
the use of a fixed, single game may look like a rather restrictive perspective and possibly
biased by the choice of the particular game. To avoid these limitations, we consider a
\textit{class} of possible games that can be played within a population
(REFERENCES). \mf{please insert some references to other papers that use a multi-game approach
  as well.} Towards this end, fix a class $\mathcal{G}$ of possible two-player symmetric games,
together with a probability measure $P_{\mathcal{G}}(G)$ for the occurrence probability of game
$G \in \mathcal{G}$. The probability measure $P_{\mathcal{G}}$ encodes the statistical
properties of the environment. A \textit{meta-game} is the tuple
$ MG=\langle CM, \mathcal{G}, P_{\mathcal{G}},F \rangle$, where $CM$ is a set of choice
mechanisms, $\mathcal{G}$ is a class of possible games, $P_{\mathcal{G}}(G)$ is the probability
of game $G$ to occur, and $F:CM \times CM \rightarrow \mathbb{R}$ is the fitness function,
defined as:
\begin{align}
  \label{eq:FittnessChoiceMechGamePairwise}
  F(c, c') = \int P_{\mathcal{G}}(G) \  F_G(c,c') \text{ d} G \,.
\end{align}
Hence, $F(c,c')$ determines the evolutionary payoff of choice mechanism $c$ against $c'$ in the
meta-game. The set $CM$ can be thought of as the set of choice mechanisms present within a given
population that is playing the games from the class $\mathcal{G}$. Consequently, it is possible
to compute the average fitness of $c$ against the population, that is given by:
\begin{align}
  \label{eq:FittnessChoiceMechGame}
  F(c) = \int P_{M}(c') \ F(c,c') \text{ d} c' = \int \int P_{\mathcal{G}}(G) \  P_{M}(c') \  F_G(c,c') \text{ d} c' \text{ d} G  \,,
\end{align}
where $P_{M}(c')$ is the probability of
encountering a co-player with choice mechanism $c'$.

Meta-games are then abstract models for the evolutionary competition between choice mechanisms
in interactive decision making contexts.  Standard notions of evolutionary game theory apply to
meta-games as well. For example, a choice mechanism $c$ is a strict Nash equilibrium if
$F(c,c) > F(c',c)$ for all $c'$; it is evolutionarily stable if for all $c'$ either (i)
$F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$ and $F(c,c') > F(c',c')$; it is neutrally stable
if for all $c'$ either (i) $F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$ and
$F(c,c') \ge F(c',c')$ \citep{Maynard-Smith1982:Evolution-and-t}. Similarly, evolutionary
dynamics can be applied to meta-games. We will later turn towards the replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St} and the replicator mutator dynamic
\citep[e.g.][]{Nowak2006:Evolutionary-Dy}.

\section{Example Application: Evolution of Preferences} \label{sec:model}

To demonstrate the potential usefulness of a meta-game approach, we will compare a small
selection of general ways of forming utility and preference representations against each
other. As for utilities, consider initially:
\begin{enumerate}
\item the \textit{objective} utility, defined by: for all $G \in \mathcal{G}$,
$$\text{obj}^G(a,a')=\pi^G(a,a');$$
\item the \textit{regret}, defined by: for all $G \in \mathcal{G}$,
$$\text{reg}^G(a,a') =\pi^G(a,a')- \max_{a''\in A} \pi^G(a'',a').$$
\end{enumerate}

@@

Although a thorough discussion of this issue goes beyond the scope of this work, let us say
that these two kinds of belief underlie two different and alternative views on
uncertainty. Faced with uncertain events, a strict Bayesian will always form a precise belief,
specified by a \textit{single} probability $\mu$. In the absence of any information about
future uncertain events, the Bayesian would then invoke the \textit{principle of insufficient
  reason}, and accordingly choose a uniform probability over the possible outcomes. In
contrast, others have argued against the obligation of representing a belief by means of a
single probability measure, opposite to the Bayesian prescriptions (REFERENCES). They argue
instead in favor of a more general account, according to which uncertainty can be unmeasurable,
and represented by a (closed and convex) set of probabilities. This line of thought has been
originated in decision theory by the famous Ellsberg's paradoxes (REF), and appears extremely
relevant in game-theoretical contexts too. Indeed, in a recent paper \citet{BattCerrMM15}
write:
